### 2016-06-03 Paper_v4.tex
A few more changes
Some initial changes on final draft

%Paragraph Edited by Mason
\subsection*{Reduced order complement network.}
The reduced order complement model described the alternate and lectin pathways (Fig. \ref{fig-schematic}).
A trigger event initiated the lectin pathway, which activated the cleavage of C2 and C4 into C2a, C2b, C4a and C4b respectively.
Classical Pathway (CP) C3 convertase (C4aC2b) then catalyzed the cleavage of C3 into C3a and C3b.
Activation of the alternative pathway was initiated through the spontaneous hydrolysis of C3 into C3a and C3b.
The C3b fragment then recombined with C3 to form the alternate pathway (AP) C3 convertase.
Both the CP and AP C3 convertases catalyzed the cleavage of C3 into C3a and C3b.
A second C3b fragment could then bind with either the CP or AP C3 convertase to form the CP (or AP) C5 convertase.
The C5 convertase catalyzed the cleavage of C5 into the C5a and C5b fragments.
Lectin pathway activation was approximated using a combination of saturation kinetics and non-linear transfer functions, which facilitated a significant reduction in the size of the model while maintaining performance.
Thus, while the reduced order complement model encoded significant biological complexity, it was highly compact consisting of only 18 differential equations and 28 model parameters.
Next, we estimated an ensemble of model parameters from time series measurements of the C3a and C5a complement proteins.

@@

%Paragraph Edited by Mason
We subsequently did a robustness analysis to interrogate the effects of perturbing initial concentrations of C3 and C5 on the levels of C3a and C5a (Fig \ref{fig-ROBUSTNESS}). We calculated robustness indices for C3a and C5a using the ratio of their area under the curves (AUCs) in the perturbed and unperturbed states for each of the 50 parameter vectors. We perturbed the initial concentration of C3, C5 and a combination of both by decreasing their concentrations by 50 \% and 90\%. We calculated corresponding robustness indices for C3a and C5a with no initiator (alternate pathway activation) and maximum initiator of lectin pathway (1g of zymosan).  A robustness index of one indicated no effect due to the perturbation whereas an index of zero indicated a strong effect of the perturbation. In both cases (50 \% and 90\% knockdown) we see that C3a and C5a levels are strongly influenced by initial concentrations of C3 and C5 in the presence of lectin pathway initiator. The C3a generated from the alternate pathway also follows the same pattern where it is influenced by the initial concentration of C3. However we observe an interesting trend with respect to C5a generated in the alternate pathway. The amount of C5a decreases with decrease in C3 concentration. This suggests that the C3b that is generated in the alternate pathway is primarily involved in the formation of AP C5-convertase (and is a rate limiting step). This is in agreement with our sensitivity analysis where were we see that the AP C3-convertase primarily can affect the dynamics of C5a formation. At the same time, in the lectin initiated pathway, the robustness of C5a levels to initial C3 concentration in both cases (robustness index of 1 or almost 1 in the 90\%) indicates that small concentrations of C3 and C5 convertases can catalyze the downstream formation of C5a. Thus the formation of C5a in presence of an initiator remains unaffected until C3 reaches a very low concentration. These results agree with the current push for complement therapeutics targeting C3 and C5. Conversely, the significance of the convertases identified by the sensitivity analysis indicate that they may also be effective targets for therapeutic intervention, however, these results require experimental validation before any definitive conclusions can be reached.
The combination of robustness analysis and sensitivity analysis, taken together, indicate specific mechanisms that can be targeted for therapeutic interventions. We summarize the sensitive mechanisms identified our model and corresponding drugs that target these mechanisms in Fig \ref{fig-targets}

@@

\subsection*{Global sensitivity analysis of model performance.}
We conducted a global sensitivity analysis to estimate which parameters controlled the performance of the reduced order model using the Sobol method \cite{SOBOL_METHOD}.
We computed the total sensitivity index of each parameter relative to the training residual for the C3a alternate, C5a alternate, C3a lectin, and C5a lectin cases.
We established the sampling bounds for each parameter from the minimum and maximum value for that parameter in the parameter ensemble.
We used the sampling method of Saltelli \textit{et al.} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
where $N$ was the number of trials, and $d$ was the number of parameters in the model \cite{saltelli2010variance}.
In our case, $N$ = 200 and $d$ = 28, so the total sensitivity indices were computed from 11,600 model evaluations.
The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \cite{SALIB}.
\clearpage

@@

\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth]{./figs/Figure4_SEN.pdf}
\caption{Sobol's sensitivity analysis of the reduced order complement model with respect to the modeling parameters.  Sensitivity analysis was conducted on the four cases we used to train our model: (A) C3a at 0 zymosan, (B) C5a 0 zymosan, (C) C3a 1 g zymosan, and (D) C5a $1$ g zymosan. The bars denote total sensitivity index which includes local contribution of each parameter and global sensitivity of significant pairwise interactions. The error bars are the 95 percent confidence interval. Pathways controlled by the sensitivity parameters (E): Bold black lines indicates the pathway is governed by one or more sensitive parameters and the red lines shows some of the current therapeutics targets.}\label{fig-SA}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth]{./figs/Figure5_RobustnessAnalysis.pdf}
\caption{Robustness analysis of the reduced order complement model with respect to the C3 and C5 initial concentrations using 50 parameter sets.  Robustness analysis was conducted on the four cases we used to train our model, C3a alternate (0 zymosan), C5a alternate (0 zymosan),  C3a lectin (1 g zymosan), and C5a lectin (1 g zymosan), by reducing the initial concentration of C3 and/or C5 by (A) 50 \% and (B) 90 \%. The bars denote robustness index which a measure of system changes from the perturbation of initial concentration that defined by the ratio of the area under the concentration curve of perturbed case and that of the unperturbed case. The error bars represent one standard deviation. At unity, the perturbed initial concentration has no impact on the measured output, and a robustness index lesser than or greater than one indicates a negative or positive relation between the perturbed initial concentration and the measured output respectively.}\label{fig-SA}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth]{./figs/Figure6_target.pdf}
\caption{ The figure graphically illustrates some of the current complement therapeutics and their targets within our network. A vast majority of the drugs target the terminal pathway by targeting C5 or C5a. Targeting C5 would reduce the formation of C5a and C5b, however targeting C5a directly would reduce the influence of the anaphylatoxin but still produce precursors for the MAC formation. A number of therapeutics also target C3 and the formation of AP C3 convertase through the inhibition of Factors B and D. Finally, very few drugs target complement initiation, this may be due to a need for balance between immunity and inflammation or other complement related diseases.  
}\label{fig-targets}
\end{figure}

>>>>

\subsection*{Reduced order complement network.}
The reduced order complement model described the alternate and lectin pathways (Fig. \ref{fig-schematic}).
A trigger event initiated the lectin pathway, which activated the cleavage of C2 and C4 into C2a, C2b, C4a and C4b respectively.
Classical Pathway (CP) C3 convertase (C4aC2b) then catalyzed the cleavage of C3 into C3a and C3b.
Activation of the alternative pathway was initiated through the spontaneous hydrolysis of C3 into C3a and C3b.
The C3b fragment then recombined with C3 to form the alternate pathway (AP) C3 convertase.
Both the CP and AP C3 convertases catalyzed the cleavage of C3 into C3a and C3b.
A second C3b fragment could then bind with either the CP or AP C3 convertase to form the CP (or AP) C5 convertase.
The C5 convertase catalyzed the cleavage of C5 into the C5a and C5b fragments.
Lectin pathway activation was approximated using a combination of saturation kinetics and non-linear transfer functions, which facilitated a significant reduction in the size of the model while maintaining performance.
Thus, while the reduced order complement model encoded significant biological complexity, it was highly compact consisting of only 18 differential equations and 28 model parameters.
Next, we estimated an ensemble of model parameters from time series measurements of the C3a and C5a complement proteins.

@@

We subsequently did a robustness analysis to interrogate the effects of perturbing initial concentrations of C3 and C5 on the levels of C3a and C5a (Fig \ref{fig-robustness-analysis}).
We calculated robustness indices for C3a and C5a using the ratio of their area under the curves (AUCs) in the perturbed and unperturbed states for each of the 50 parameter vectors. We perturbed the initial concentration of C3, C5 and a combination of both by decreasing their concentrations by 50 \% and 90\%. We calculated corresponding robustness indices for C3a and C5a with no initiator (alternate pathway activation) and maximum initiator of lectin pathway (1g of zymosan).  A robustness index of one indicated no effect due to the perturbation whereas an index of zero indicated a strong effect of the perturbation. In both cases (50 \% and 90\% knockdown) we see that C3a and C5a levels are strongly influenced by initial concentrations of C3 and C5 in the presence of lectin pathway initiator. The C3a generated from the alternate pathway also follows the same pattern where it is influenced by the initial concentration of C3. However we observe an interesting trend with respect to C5a generated in the alternate pathway. The amount of C5a decreases with decrease in C3 concentration. This suggests that the C3b that is generated in the alternate pathway is primarily involved in the formation of AP C5-convertase (and is a rate limiting step). This is in agreement with our sensitivity analysis where were we see that the AP C3-convertase primarily can affect the dynamics of C5a formation. At the same time, in the lectin initiated pathway, the robustness of C5a levels to initial C3 concentration in both cases (robustness index of 1 or almost 1 in the 90\%) indicates that small concentrations of C3 and C5 convertases can catalyze the downstream formation of C5a. Thus the formation of C5a in presence of an initiator remains unaffected until C3 reaches a very low concentration. These results agree with the current push for complement therapeutics targeting C3 and C5. Conversely, the significance of the convertases identified by the sensitivity analysis indicate that they may also be effective targets for therapeutic intervention, however, these results require experimental validation before any definitive conclusions can be reached.
The combination of robustness analysis and sensitivity analysis, taken together, indicate specific mechanisms that can be targeted for therapeutic interventions.

%We summarize the sensitive mechanisms identified our model and corresponding drugs that target these mechanisms in Fig \ref{fig-targets}

@@

\subsection*{Sensitivity and robustness analysis of model performance.}
We conducted a global sensitivity analysis to estimate which parameters controlled the performance of the reduced order model using the Sobol method \cite{SOBOL_METHOD}.
We computed the total sensitivity index of each parameter relative to the training residual for the C3a alternate, C5a alternate, C3a lectin, and C5a lectin cases.
We established the sampling bounds for each parameter from the minimum and maximum value for that parameter in the parameter ensemble.
We used the sampling method of Saltelli \textit{et al.} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
where $N$ was the number of trials, and $d$ was the number of parameters in the model \cite{saltelli2010variance}.
In our case, $N$ = 200 and $d$ = 28, so the total sensitivity indices were computed from 11,600 model evaluations.
The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \cite{SALIB}.

Robustness coefficients were calculated as shown previously \cite{Tasseff:2011aa}.
Robustness coefficients quantify the response of a marker to a structural or operational perturbation to the network architecture.
Robustness coefficients denoted by $\alpha\left(i,j,t_{o},t_{f}\right)$ are defined as:
\begin{equation}
\alpha\left(i,j,t_{o},t_{f}\right)=\left(\displaystyle\int_{t_{o}}^{t_{f}}x_{i}\left(t\right)dt\right)^{-1}\left(\displaystyle\int_{t_{o}}^{t_{f}}x^{(j)}_{i}\left(t\right)dt\right)
\end{equation}
Here $t_o$ and $t_f$ denote the initial and final simulation time respectively, while $i$ and $j$ denote the indices for the marker and the perturbation respectively.
A value of $\alpha\left(i,j,t_{o},t_{f}\right)>1$, indicates increased marker abundance, while $\alpha\left(i,j,t_{o},t_{f}\right)<1$ indicates decreased marker abundance following perturbation $j$. If $\alpha\left(i,j,t_{o},t_{f}\right)\sim{1}$ the $j$th perturbation does not influence the abundance of marker $i$.

@@

\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth]{./figs/Figure4_SEN_v2.pdf}
\caption{Sobol's sensitivity analysis of the reduced order complement model with respect to the modeling parameters.  Sensitivity analysis was conducted on the four cases we used to train our model: (A) C3a at 0g zymosan, (B) C5a 0g zymosan, (C) C3a 1g zymosan, and (D) C5a 1g zymosan. The bars denote total sensitivity index which includes local contribution of each parameter and global sensitivity of significant pairwise interactions. The error bars are the 95 percent confidence interval. Pathways controlled by the sensitivity parameters (E): Bold black lines indicates the pathway is governed by one or more sensitive parameters and the red lines shows some of the current therapeutics targets. Red indicates current complement therapeutics. }\label{fig-SA}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth]{./figs/Figure5_RobustnessAnalysis.pdf}
\caption{Robustness analysis of the reduced order complement model with respect to the C3 and C5 initial concentrations using 50 parameter sets.  Robustness analysis was conducted on the four cases we used to train our model, C3a alternate (0 zymosan), C5a alternate (0 zymosan),  C3a lectin (1 g zymosan), and C5a lectin (1 g zymosan), by reducing the initial concentration of C3 and/or C5 by (A) 50 \% and (B) 90 \%. The bars denote robustness index which a measure of system changes from the perturbation of initial concentration that defined by the ratio of the area under the concentration curve of perturbed case and that of the unperturbed case. The error bars represent one standard deviation. At unity, the perturbed initial concentration has no impact on the measured output, and a robustness index lesser than or greater than one indicates a negative or positive relation between the perturbed initial concentration and the measured output respectively.}\label{fig-robustness-analysis}
\end{figure}

% \begin{figure}[h]
% \centering
% \includegraphics[width=1.0\textwidth]{./figs/Figure6_target.pdf}
% \caption{ The figure graphically illustrates some of the current complement therapeutics and their targets within our network. A vast majority of the drugs target the terminal pathway by targeting C5 or C5a. Targeting C5 would reduce the formation of C5a and C5b, however targeting C5a directly would reduce the influence of the anaphylatoxin but still produce precursors for the MAC formation. A number of therapeutics also target C3 and the formation of AP C3 convertase through the inhibition of Factors B and D. Finally, very few drugs target complement initiation, this may be due to a need for balance between immunity and inflammation or other complement related diseases.
% }\label{fig-targets}
% \end{figure}
